---
title: "Compression of Paths"
author: "Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compression of Paths}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

See the [Working with File Paths in R Scripts](introduction.html) for an
introduction on how we suggest to handle file paths in R scripts. With this
vignette I want to recapitulate what the functions do that somehow deal with
path dictionaries and that I extracted from the package kwb.fakin into this
package.

## Provide Example File Paths

We start by loading some example `paths`:

```{r}
# Read and print example (file) paths
writeLines(paths <- kwb.pathdict:::example_paths())

# Manually define some (folder) paths
directories <- c(
  "short/path/to/directory", 
  "short/path/to/directory", 
  "short/path/to/directory", 
  "longer/path/to/another/directory", 
  "longer/path/to/another/directory"
)
``` 

## Compression of Folder Paths

Assume you have a folder with a long path (either due to a lot of levels or due
to long folder names). This long path is repeated for each file in the folder. 
The package kwb.pathdict provides a `compress()` function that may be used to
introduce short names for directory paths.

```{r}
short_dirs <- kwb.pathdict:::compress(directories)

as.character(short_dirs)
```

```{r}
attributes(kwb.pathdict:::compress_one_by_one(paths, n = 3))
```

## Functions compress_paths() and sorted_importance()

The function `compress_paths()` takes a vector of full file paths as input. The
paths are "compressed" by replacing all directory paths by placeholders \<a1\>,
\<a2\>, \<a3\> as shown in the following:

```{r}
# Compress the paths
paths_1 <- kwb.pathdict:::compress_paths(paths)

# Show the compressed paths, the dictionary and the total size
kwb.pathdict:::print_compressed(paths_1)
```

Equal directories are replaced by equal placeholders. The placeholders and their
"values" are stored in a so called dictionary. The dictionary can be seen as a
lookup table that maps short names to long names. The dictionary is returned
in the attribute `dict` of what `compress_paths()` returns:

```{r}
(dictionary_1 <- kwb.utils::getAttribute(paths_1, "dict"))
```

The dictionary is sorted by the importance of the strings (paths) that have been
given to `compress()` so that the most important paths get the highest positions 
in the dictionary. 

The placeholder \<a1\> is given to the most "important" directory path, the
placeholder \<a2\> to the second most "important", and so on.

With "importance" we mean the product of the frequency of the directory path
(its number of occurrences) and its string length. It is internally calculated
by the function `sorted_importance()`. This function returns the product of
string frequency and string length for the different strings given in its first
argument.

```{r}
kwb.pathdict:::named_vector_to_data_frame(
  kwb.pathdict:::sorted_importance(dirname(paths)
  )
)
```

For example, the importance of the path ``r (d1 <- (d <- dirname(paths))[1])``
is `r kwb.pathdict:::sorted_importance(d)[d1]` because it is `r nchar(d1)`
characters long and it appears `r table(d)[d1]` times in the vector `paths`:

```{r}
# Paths of directories
dir_paths <- dirname(paths)

# Frequency of the first path
(path_frequency <- unname(table(dir_paths)[dir_paths[1]]))

# Length of the first path (in number of characters)
(path_length <- nchar(dir_paths[1]))

# Importance of the first path
path_frequency * path_length
```

Just as the original file paths contain common directory paths, the paths in the
directory also contain common base paths. These common sub-paths can themselves
be replaced by further placeholders. Therefore, we set the argument `maxdepth`
of `compress_paths()`:

```{r}
# Compress the full file paths by replacing directory paths by placeholders
paths_2 <- kwb.pathdict:::compress_paths(paths, maxdepth = 2)

# Show the short paths (without attribute "dict")
structure(paths_2, dict = NULL)

# The new paths are identical to the paths created with maxdepth = 1
identical(structure(paths_2, dict = NULL), structure(paths_1, dict = NULL))

# However, the "dictionary" paths are themselves shortened by new placeholders
kwb.pathdict:::print_dict(compressed = paths_2)
```

The following code section demonstrates how increasing `maxdepth` step by step compresses the dictionary even further:

```{r}
kwb.pathdict:::print_dict(paths, 3)
kwb.pathdict:::print_dict(paths, 4)
kwb.pathdict:::print_dict(paths, 5)
```

The dictionary is required to convert the short names back to the original
paths using the `resolve()` function of the `kwb.utils` package:

```{r collapse = TRUE}
# Reproduce the original file paths from the compressed paths and the dictionary
(reproduced_paths <- kwb.utils::resolve(paths_1, dictionary_1))

# Check if we really get back the original paths
identical(paths_1, reproduced_paths)
```

Ok, I understand what happened but what can I use this for? The idea was to 
compress very big path lists in an effective but comprehensive and 
human-readable way. If we compare the size of the original path vector with
the size of the compressed path vector including the dictionary, we see that the
idea does not work at all for the given example paths:

```{r}
# Object sizes in bytes for maxdepth = 1 to 5
sapply(1:5, kwb.pathdict:::dictionary_size, x = paths)
```

However, using big file lists with thousands of paths, we get a completely
different picture:

```{r}
file_info <- fs::dir_info("~/Downloads/kwb", recursive = TRUE, type = "file")
many_paths <- as.character(file_info$path)

# Size of path vector
object.size(many_paths)

# Size of compressed path vectors including dictionary
kwb.pathdict:::dictionary_size(many_paths, 1)
kwb.pathdict:::dictionary_size(many_paths, 2)
kwb.pathdict:::dictionary_size(many_paths, 3)
kwb.pathdict:::dictionary_size(many_paths, 4)
kwb.pathdict:::dictionary_size(many_paths, 5)

# How do the dictionaries look like?
kwb.pathdict:::print_dict(many_paths, 1)
kwb.pathdict:::print_dict(many_paths, 2)
kwb.pathdict:::print_dict(many_paths, 3)
kwb.pathdict:::print_dict(many_paths, 4)
kwb.pathdict:::print_dict(many_paths, 5)
```

There seems to be a problem. There are unexpected duplicates!

```{r}
for (m in 1:5) {
  message("m = ", m)
  kwb.pathdict:::print_dict(m = m, c(
    "a/b/c/d/e1/file1.txt", 
    "a/b/c/d/e2/file2.txt"
  ))
}
```

The function `compress_paths()` is a high level function that internally calls
the helper function `compress()`. That function is demonstrated next.

### Function compress()

The function `compress()` is a helper function that is called internally by the
function `compress_paths()`. It replaces all distinct values in a vector with a
short term that ist formatted as a "placeholder". It returns a "dictionary",
i.e. a list that contains the original values as values and the short terms as
keys

```{r}
compressed <- kwb.pathdict:::compress(c("abc", "abc", "defghi"))

kwb.utils::getAttribute(compressed, "dict")

x1 <- c(rep("short", 1), rep("very much longer", 1))
x2 <- c(rep("short", 5), rep("very much longer", 1))

kwb.utils::getAttribute(kwb.pathdict:::compress(x1), "dict")
kwb.utils::getAttribute(kwb.pathdict:::compress(x2), "dict")
```

The elements in the dictionary are ordered by "importance", i.e. the product
of frequency and length of the strings (see function `sorted_importance()` 
above).

There are three more compress functions that are described next.

### compress_one_by_one()

What does this function do?

```{r}
out <- capture.output(result <- kwb.pathdict:::compress_one_by_one(x1))
result
```

### compress_with_dictionary()

The `compress_with_dictionary()` function expects a matrix of character and a
dictionary, i.e. a list of values that are assigned to short keywords as input.
Values that are contained in the dictionary are replaced by their short key. 
Values that are not contained are replaced with what is given in `fill.value`.
But why? I think that this is just to treat empty values in the matrix. For all
non-empty values it is assumed that a corresponding entry is contained in the
dictionary.

```{r}
(subdirs_1 <- matrix(c("abc", "def", "ghi", "jkl"), 2, 2))
(subdirs_2 <- matrix(c("abc", "def", "ghi", ""), 2, 2))

kwb.pathdict:::compress_with_dictionary(
  subdirs_1,
  dict = list(a = "abc", d = "def", g = "ghi", k = "klm"),
  fill.value = "_filled_"
)

kwb.pathdict:::compress_with_dictionary(
  subdirs_2,
  dict = list(a = "abc", d = "def", g = "ghi", j = "jkl"),
  fill.value = "_filled_"
)
```
