---
title: "Compression of Paths"
author: "Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compression of Paths}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

See the [Working with File Paths in R Scripts](introduction.html) for an 
introduction on how we suggest to handle file paths in R scripts.

## Define some Example File Paths

```{r}
paths <- c(
  "//very/long/path/to/the/projects/project-1/wp-1/input/file 1.csv",
  "//very/long/path/to/the/projects/project-1/wp-1/input/file-2.csv",
  "//very/long/path/to/the/projects/project-1/wp-1/analysis/summary.pdf",
  "//very/long/path/to/the/projects/project-1/wp 2/input/köpenick_dirty.csv",
  "//very/long/path/to/the/projects/project-1/wp 2/output/koepenick_clean.csv",
  "//very/long/path/to/the/projects/project-2/Daten/file-1.csv",
  "//very/long/path/to/the/projects/project-2/Grafiken/file 1.png",
  "//very/long/path/to/the/projects/project-2/Berichte/bericht-1.doc",
  "//very/long/path/to/the/projects/project-2/Berichte/bericht-2.doc"
)
``` 

## Compress the Paths

```{r}
kwb.pathdict:::compress_paths(paths)
kwb.pathdict:::compress_paths(paths, maxdepth = 2)
kwb.pathdict:::compress_paths(paths, maxdepth = 4)
```

## Importance of Directories

In order to find the most important folder paths in terms of frequency and
length you may use the function `sorted_importance()`:

```{r}
kwb.pathdict:::sorted_importance(dirname(paths))
```

## Compression of Folder Paths

Assume you have a folder with a long path (either due to a lot of levels or due
to long folder names). This long path is repeated for each file in the folder. 
The package kwb.pathdict provides a `compress()` function that may be used to
introduce short names for directory paths.

```{r}
directories <- c(
  "short/path/to/directory", 
  "short/path/to/directory", 
  "short/path/to/directory", 
  "longer/path/to/another/directory", 
  "longer/path/to/another/directory"
)

short_dirs <- kwb.pathdict:::compress(directories)

as.character(short_dirs)
```

The function returns the lookup table (dictionary) that maps short names to long 
names in the attribute `dict`:

```{r collapse = TRUE}
(dictionary <- kwb.utils::getAttribute(short_dirs, "dict"))
```

The dictionary is sorted by the importance of the strings (paths) that have been
given to `compress()` so that the most important paths get the highest positions 
in the dictionary. With importance we mean the product of the length of a string 
(path) and its number of occurrences.

The dictionary is required to convert the short names back to the original
paths using the `resolve()` function of the `kwb.utils` package:

```{r collapse = TRUE}
(long_dirs <- kwb.utils::resolve(short_dirs, dictionary))

identical(long_dirs, directories)
```

There is another function `compress_one_by_one`, what does it do?

```{r}
attributes(kwb.pathdict:::compress_one_by_one(paths, n = 3))
```
