---
title: "Compression of Paths"
author: "Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compression of Paths}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

See the [Working with File Paths in R Scripts](introduction.html) for an
introduction on how we suggest to handle file paths in R scripts. With this
vignette I want to recapitulate what the functions do that somehow deal with
path dictionaries and that I extracted from the package kwb.fakin into this
package.

We start by loading some example `paths`:

```{r}
writeLines(paths <- kwb.pathdict:::example_paths())
``` 

## Functions compress_paths() and sorted_importance()

The function `compress_paths()` takes a vector of full file paths as input. The
paths are "compressed" by replacing all directory paths by placeholders \<a1\>,
\<a2\>, \<a3\> as shown in the following:

```{r}
# Compress the paths
paths_1 <- kwb.pathdict:::compress_paths(paths)

# Show the compressed paths, the dictionary and the total size
kwb.pathdict:::print_compressed(paths_1)
```

Equal directories are replaced by equal placeholders. The placeholders and their
"values" are stored in a so called dictionary. The placeholder \<a1\> is given
to the most "important" directory path, the placeholder \<a2\> to the second
most "important", and so on. With "importance" we mean the product of the
frequency of the directory path and its string length. It is internally
calculated by the function `sorted_importance()`:

```{r}
kwb.pathdict:::named_vector_to_data_frame(
  kwb.pathdict:::sorted_importance(dirname(paths)
  )
)
```

For example, the importance of the path ``r (d1 <- (d <- dirname(paths))[1])``
is `r kwb.pathdict:::sorted_importance(d)[d1]` because it is `r nchar(d1)`
characters long and it appears `r table(d)[d1]` times in the vector `paths`:

```{r}
# Paths of directories
dir_paths <- dirname(paths)

# Frequency of the first path
(path_frequency <- unname(table(dir_paths)[dir_paths[1]]))

# Length of the first path (in number of characters)
(path_length <- nchar(dir_paths[1]))

# Importance of the first path
path_frequency * path_length
```

Just as the original file paths contain common directory paths, the paths in the
directory also contain common base paths. These common sub-paths can themselves
be replaced by further placeholders. Therefore, we set the argument `maxdepth`
of `compress_paths()`:

```{r}
# Compress the full file paths by replacing directory paths by placeholders
paths_2 <- kwb.pathdict:::compress_paths(paths, maxdepth = 2)

# Show the short paths (without attribute "dict")
structure(paths_2, dict = NULL)

# The new paths are identical to the paths created with maxdepth = 1
identical(structure(paths_2, dict = NULL), structure(paths_1, dict = NULL))

# However, the "dictionary" paths are themselves shortened by new placeholders
kwb.pathdict:::print_dict(compressed = paths_2)
```

The following code section demonstrates how increasing `maxdepth` step by step compresses the dictionary even further:

```{r}
kwb.pathdict:::print_dict(paths, 3)
kwb.pathdict:::print_dict(paths, 4)
kwb.pathdict:::print_dict(paths, 5)
```

Ok, I understand what happened but what can I use this for? The idea was to 
compress very big path lists in an effective but comprehensive and 
human-readable way. If we compare the size of the original path vector with
the size of the compressed path vector including the dictionary, we see that the
idea does not work at all for the given example paths:

```{r}
# Object sizes in bytes for maxdepth = 1 to 5
sapply(1:5, kwb.pathdict:::dictionary_size, x = paths)
```

However, using big file lists with thousands of paths, we get a completely
different picture:

```{r}
file_info <- fs::dir_info("~/Downloads/kwb", recursive = TRUE, type = "file")
many_paths <- as.character(file_info$path)

# Size of path vector
object.size(many_paths)

# Size of compressed path vectors including dictionary
kwb.pathdict:::dictionary_size(many_paths, 1)
kwb.pathdict:::dictionary_size(many_paths, 2)
kwb.pathdict:::dictionary_size(many_paths, 3)
kwb.pathdict:::dictionary_size(many_paths, 4)
kwb.pathdict:::dictionary_size(many_paths, 5)

# How do the dictionaries look like?
kwb.pathdict:::print_dict(many_paths, 1)
kwb.pathdict:::print_dict(many_paths, 2)
kwb.pathdict:::print_dict(many_paths, 3)
kwb.pathdict:::print_dict(many_paths, 4)
kwb.pathdict:::print_dict(many_paths, 5)
```

There seems to be a problem. There are unexpected duplicates!

```{r}
for (m in 1:5) {
  message("m = ", m)
  kwb.pathdict:::print_dict(m = m, c(
    "a/b/c/d/e1/file1.txt", 
    "a/b/c/d/e2/file2.txt"
  ))
}
```

The function `compress_paths()` is a high level function that internally calls
the helper function `compress()`. That function is demonstrated next.

### Function compress()

The function `compress()` is a helper function that is called internally by the
function `compress_paths()`. It replaces all distinct values in a vector with a
short term that ist formatted as a "placeholder". It returns a "dictionary",
i.e. a list that contains the original values as values and the short terms as
keys

```{r}
compressed <- kwb.pathdict:::compress(c("abc", "abc", "defghi"))

kwb.utils::getAttribute(compressed, "dict")

x1 <- c(rep("short", 1), rep("very much longer", 1))
x2 <- c(rep("short", 5), rep("very much longer", 1))

kwb.utils::getAttribute(kwb.pathdict:::compress(x1), "dict")
kwb.utils::getAttribute(kwb.pathdict:::compress(x2), "dict")
```

The elements in the dictionary are ordered by "importance", i.e. the product
of frequency and length of the strings (see next).

### sorted_importance()

This function returns the product of string frequency and string length for the
different strings given in x.

```{r}
x1
kwb.pathdict:::sorted_importance(x1)

x2
kwb.pathdict:::sorted_importance(x2)
```

There are three more compress functions.

### compress_one_by_one()

```{r}
out <- capture.output(result <- kwb.pathdict:::compress_one_by_one(x1))
result
```

### compress_with_dictionary()

The `compress_with_dictionary()` function expects a matrix of character and a
dictionary, i.e. a list of values that are assigned to short keywords as input.
Values that are contained in the dictionary are replaced by their short key. 
Values that are not contained are replaced with what is given in `fill.value`.
But why? I think that this is just to treat empty values in the matrix. For all
non-empty values it is assumed that a corresponding entry is contained in the
dictionary.

```{r}
(subdirs_1 <- matrix(c("abc", "def", "ghi", "jkl"), 2, 2))
(subdirs_2 <- matrix(c("abc", "def", "ghi", ""), 2, 2))

kwb.pathdict:::compress_with_dictionary(
  subdirs_1,
  dict = list(a = "abc", d = "def", g = "ghi", k = "klm"),
  fill.value = "_filled_"
)

kwb.pathdict:::compress_with_dictionary(
  subdirs_2,
  dict = list(a = "abc", d = "def", g = "ghi", j = "jkl"),
  fill.value = "_filled_"
)
```

### data_frame_to_paths()

The next function, `data_frame_to_paths()` converts a subdirectory matrix that
has been converted to a data frame back to the paths by pasting all non-empty
columns with slash. Short paths must have empty values `""` and not `NA` in the
last columns:

```{r}
(df <- kwb.utils::asNoFactorDataFrame(matrix(
  c("a", "b", "c", 
    "d", "e", "",
    "f", "g", NA), 3, 3, byrow = TRUE
)))

kwb.pathdict:::data_frame_to_paths(df)
```

```{r}
kwb.pathdict:::get_dictionary_one_by_one
kwb.pathdict:::get_frequency_score
kwb.pathdict:::get_next_level
kwb.pathdict:::get_subdir_frequencies
kwb.pathdict:::get_subdirs_by_frequence
kwb.pathdict:::is_placeholder
kwb.pathdict:::log_result_if
kwb.pathdict:::lookup_in_dictionary
kwb.pathdict:::main_columns_winner
kwb.pathdict:::print_path_frequencies
kwb.pathdict:::replace_subdirs
kwb.pathdict:::rescore_and_reorder_frequency_data

kwb.pathdict:::starts_with_parts
kwb.pathdict:::to_dictionary
kwb.pathdict:::to_dictionary_key
kwb.pathdict:::to_frequency_data
kwb.pathdict:::to_placeholder
kwb.pathdict:::update_frequency_data_length
kwb.pathdict:::use_dictionary
```


## Define some Example File Paths

```{r}
paths <- c(
  "//very/long/path/to/the/projects/project-1/wp-1/input/file 1.csv",
  "//very/long/path/to/the/projects/project-1/wp-1/input/file-2.csv",
  "//very/long/path/to/the/projects/project-1/wp-1/analysis/summary.pdf",
  "//very/long/path/to/the/projects/project-1/wp 2/input/köpenick_dirty.csv",
  "//very/long/path/to/the/projects/project-1/wp 2/output/koepenick_clean.csv",
  "//very/long/path/to/the/projects/project-2/Daten/file-1.csv",
  "//very/long/path/to/the/projects/project-2/Grafiken/file 1.png",
  "//very/long/path/to/the/projects/project-2/Berichte/bericht-1.doc",
  "//very/long/path/to/the/projects/project-2/Berichte/bericht-2.doc"
)
``` 

## Compress the Paths

```{r}
kwb.pathdict:::compress_paths(paths)
kwb.pathdict:::compress_paths(paths, maxdepth = 2)
kwb.pathdict:::compress_paths(paths, maxdepth = 4)
```

## Importance of Directories

In order to find the most important folder paths in terms of frequency and
length you may use the function `sorted_importance()`:

```{r}
kwb.pathdict:::sorted_importance(dirname(paths))
```

## Compression of Folder Paths

Assume you have a folder with a long path (either due to a lot of levels or due
to long folder names). This long path is repeated for each file in the folder. 
The package kwb.pathdict provides a `compress()` function that may be used to
introduce short names for directory paths.

```{r}
directories <- c(
  "short/path/to/directory", 
  "short/path/to/directory", 
  "short/path/to/directory", 
  "longer/path/to/another/directory", 
  "longer/path/to/another/directory"
)

short_dirs <- kwb.pathdict:::compress(directories)

as.character(short_dirs)
```

The function returns the lookup table (dictionary) that maps short names to long 
names in the attribute `dict`:

```{r collapse = TRUE}
(dictionary <- kwb.utils::getAttribute(short_dirs, "dict"))
```

The dictionary is sorted by the importance of the strings (paths) that have been
given to `compress()` so that the most important paths get the highest positions 
in the dictionary. With importance we mean the product of the length of a string 
(path) and its number of occurrences.

The dictionary is required to convert the short names back to the original
paths using the `resolve()` function of the `kwb.utils` package:

```{r collapse = TRUE}
(long_dirs <- kwb.utils::resolve(short_dirs, dictionary))

identical(long_dirs, directories)
```

There is another function `compress_one_by_one`, what does it do?

```{r}
attributes(kwb.pathdict:::compress_one_by_one(paths, n = 3))
```
