---
title: "Functions in this Package"
author: "Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functions}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette describes the functions that have not yet been described in other
vignettes. 

### Provide Random Example Paths

```{r}
set.seed(53653815)

(paths <- kwb.pathdict::random_paths(5, min_chars = 5, max_elements = 4))
```

### data_frame_to_paths()

The function `data_frame_to_paths()` converts a subdirectory matrix that has
been converted to a data frame back to the paths by pasting all non-empty
columns with slashes. Paths that do not have the maximum depth must have empty 
values `""` (and not `NA`) in the last columns, otherwise the string `NA` will
appear in the paths.

```{r}
# Create subdirectory matrix
subdir_matrix <- kwb.file::to_subdir_matrix(paths)

# Convert to data frame
(subdir_data <- kwb.utils::asNoFactorDataFrame(subdir_matrix))

# Reconstruct the paths from this data frame
paths_reconstructed <- kwb.pathdict:::data_frame_to_paths(subdir_data)

# Check that the paths have not changed
identical(paths_reconstructed, paths)
```

### is_placeholder

This function checks if a string looks like a \<placeholder\>:

```{r}
kwb.pathdict:::is_placeholder("abc")
kwb.pathdict:::is_placeholder("<abc>")
kwb.pathdict:::is_placeholder(c("<abc>", "<def>", "ghi", "<jkl>"))
kwb.pathdict:::is_placeholder("<abc>def")
kwb.pathdict:::is_placeholder("<abc><def>")
kwb.pathdict:::is_placeholder("<>")
kwb.pathdict:::is_placeholder("[abc]")
```

### get_dictionary_one_by_one

```{r}
result <- kwb.pathdict:::get_dictionary_one_by_one(paths, n = 5)
result
```

### Subfolder Frequency Functions

The function `to_frequency_data()` creates a data frame from a list with paths
as keys and path frequencies as values.

```{r}
# Define artificial frequency data
freqs <- list(a = 10, bc = 20, cde = 15)

# Convert to data frame, calculate string length and score
(frequency_data <- kwb.pathdict:::to_frequency_data(freqs))
```

The function `rescore_and_reorder_frequency_data()` takes a data frame with
columns `length` and `count` as input. It calculates `score = (length -
nchar(to_placeholder(key))) * count` and orders the data frame decreasingly by
this score.

```{r}
# Define artificial frequency data
key <- "abc"

kwb.pathdict:::rescore_and_reorder_frequency_data(frequency_data, key)
```

This function internally uses `get_frequency_score()`:

```{r}
kwb.pathdict:::get_frequency_score(frequency_data, key)

# Calculate manually
(frequency_data$length - nchar(key) - 2) * frequency_data$count
```

### get_subdir_frequencies()

The function `get_subdir_frequencies()` either gets a vector of path strings or
a list of character vectors representing all parts of the full paths as input.
For each possible path length it determines all different sub-paths and orders
them by their "importance", i.e. the product of frequency and path length in
number of characters. 

```{r}
length(paths)

kwb.pathdict:::get_subdir_frequencies(
  paths = paths, first.only = FALSE, dbg = FALSE
)
```

### get_next_level()

The function `get_next_level()` is a helper function that is called by the
function `compress_one_by_one()`.

```{r}
kwb.pathdict:::get_next_level
```

### Functions that still need to be described

```{r}
kwb.pathdict:::get_subdirs_by_frequence
kwb.pathdict:::log_result_if
kwb.pathdict:::lookup_in_dictionary
kwb.pathdict:::main_columns_winner
kwb.pathdict:::print_path_frequencies
kwb.pathdict:::replace_subdirs
kwb.pathdict:::starts_with_parts
kwb.pathdict:::to_dictionary
kwb.pathdict:::to_dictionary_key
kwb.pathdict:::to_placeholder
kwb.pathdict:::update_frequency_data_length
kwb.pathdict:::use_dictionary
```
