---
title: "Functions in this Package"
author: "Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette aims at describing the functions that have not yet been described
in other vignettes, one by one. For the moment I only print out the 
implementation of each function.

### data_frame_to_paths()

The next function, `data_frame_to_paths()` converts a subdirectory matrix that
has been converted to a data frame back to the paths by pasting all non-empty
columns with slash. Short paths must have empty values `""` and not `NA` in the
last columns:

```{r}
(df <- kwb.utils::asNoFactorDataFrame(matrix(
  c("a", "b", "c", 
    "d", "e", "",
    "f", "g", NA), 3, 3, byrow = TRUE
)))

kwb.pathdict:::data_frame_to_paths(df)
```

### Functions that still need to be des described

```{r}
kwb.pathdict:::get_dictionary_one_by_one
kwb.pathdict:::get_frequency_score
kwb.pathdict:::get_next_level
kwb.pathdict:::get_subdir_frequencies
kwb.pathdict:::get_subdirs_by_frequence
kwb.pathdict:::is_placeholder
kwb.pathdict:::log_result_if
kwb.pathdict:::lookup_in_dictionary
kwb.pathdict:::main_columns_winner
kwb.pathdict:::print_path_frequencies
kwb.pathdict:::replace_subdirs
kwb.pathdict:::rescore_and_reorder_frequency_data
kwb.pathdict:::starts_with_parts
kwb.pathdict:::to_dictionary
kwb.pathdict:::to_dictionary_key
kwb.pathdict:::to_frequency_data
kwb.pathdict:::to_placeholder
kwb.pathdict:::update_frequency_data_length
kwb.pathdict:::use_dictionary
```
