---
title: "Functions in this Package"
author: "Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functions}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette describes the functions that have not yet been described in other
vignettes. 

### Provide Random Example Paths

```{r}
set.seed(53653815)

(paths <- kwb.pathdict::random_paths(4, min_chars = 5, max_elements = 3))
```

### data_frame_to_paths()

The function `data_frame_to_paths()` converts a subdirectory matrix that has
been converted to a data frame back to the paths by pasting all non-empty
columns with slashes. Paths that do not have the maximum depth must have empty 
values `""` (and not `NA`) in the last columns, otherwise the string `NA` will
appear in the paths.

```{r}
# Create subdirectory matrix as data frame
(subdir_data <- kwb.file::to_subdir_matrix(paths, result_type = "data.frame"))

# Reconstruct the paths from this data frame
paths_reconstructed <- kwb.pathdict:::data_frame_to_paths(subdir_data)

# Check that the paths have not changed
identical(paths_reconstructed, paths)
```

### is_placeholder

This function checks if a string looks like a \<placeholder\>:

```{r}
kwb.pathdict:::is_placeholder("abc")
kwb.pathdict:::is_placeholder("<abc>")
kwb.pathdict:::is_placeholder(c("<abc>", "<def>", "ghi", "<jkl>"))
kwb.pathdict:::is_placeholder("<abc>def")
kwb.pathdict:::is_placeholder("<abc><def>")
kwb.pathdict:::is_placeholder("<>")
kwb.pathdict:::is_placeholder("[abc]")
```

### get_dictionary_one_by_one

```{r}
result <- kwb.pathdict:::get_dictionary_one_by_one(paths, n = 5)
result
```

### Subfolder Frequency Functions

At the start of `get_dictionary_one_by_one()` the function
`get_subdir_frequencies()` is called. This function can be given a vector of
path strings as input. For each possible depth in the path tree it determines
all different sub-paths and orders them by their "importance", i.e. the product
of frequency and path length in number of characters. By default, only the most
important path per depth level is returned:

```{r}
frequencies <- kwb.pathdict:::get_subdir_frequencies(paths = paths, dbg = FALSE)

# Define helper function for output
print_frequencies <- function(f) {
  for (i in seq_along(f)) {
    message("\ndepth = ", i)
    print(kwb.pathdict:::named_vector_to_data_frame(f[[i]]))
  }
}

print_frequencies(frequencies)
```

All different paths are returned when the argument `first.only` is set to 
`FALSE`:

```{r}
all_frequencies <- kwb.pathdict:::get_subdir_frequencies(
  paths = paths, first.only = FALSE, dbg = FALSE
)

print_frequencies(all_frequencies)
```

The function `to_frequency_data()` is called next within 
`get_dictionary_one_by_one()`. It takes the list returned by 
`get_subdir_frequencies()` as input and converts it to a data frame.

```{r}
# Convert frequency list to data frame
(frequency_data <- kwb.pathdict:::to_frequency_data(frequencies))
(all_frequency_data <- kwb.pathdict:::to_frequency_data(all_frequencies))
```

The function `rescore_and_reorder_frequency_data()` takes a data frame with
columns `length` and `count` as input. It calculates `score = (length -
nchar(to_placeholder(key))) * count` and orders the data frame decreasingly by
this score.

```{r}
# Define artificial frequency data
key <- "abc"

kwb.pathdict:::rescore_and_reorder_frequency_data(frequency_data, key)
```

### get_next_level()

The function `get_next_level()` is a helper function that is called by the
function `compress_one_by_one()`.

```{r}
kwb.pathdict:::get_next_level
```

### get_subdirs_by_frequence()

The function `get_subdirs_by_frequence()` is used by a FAKIN script that is not
yet part of a package. It gets different inputs:

1. `subdirs`: subdirectory matrix
2. `cumid`: cumulative identifier, created with
  `kwb.pathdict:::to_cumulative_id()`
3. `freqinfo`: one-row data frame with columns `depth`, `n.x`, `n.Freq`

```{r}
subdirs <- kwb.file::to_subdir_matrix(paths)
cumid <- kwb.pathdict:::to_cumulative_id(subdirs)
freqinfo <- data.frame(depth = 2, n.x = 10, n.Freq = 10)

kwb.pathdict:::get_subdirs_by_frequence(subdirs, cumid, freqinfo)
```

### Functions that still need to be described

```{r}
kwb.pathdict:::log_result_if
kwb.pathdict:::lookup_in_dictionary
kwb.pathdict:::main_columns_winner
kwb.pathdict:::print_path_frequencies
kwb.pathdict:::replace_subdirs
kwb.pathdict:::starts_with_parts
kwb.pathdict:::to_dictionary
kwb.pathdict:::to_dictionary_key
kwb.pathdict:::to_placeholder
kwb.pathdict:::update_frequency_data_length
kwb.pathdict:::use_dictionary
```
