---
title: "Functions in this Package"
author: "Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functions}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette describes the functions that have not yet been described in other
vignettes. 

### Provide Random Example Paths

```{r}
set.seed(53653815)

(paths <- kwb.pathdict::random_paths(4, min_chars = 5, max_elements = 3))
```

### data_frame_to_paths()

The function `data_frame_to_paths()` converts a subdirectory matrix that has
been converted to a data frame back to the paths by pasting all non-empty
columns with slashes. Paths that do not have the maximum depth must have empty 
values `""` (and not `NA`) in the last columns, otherwise the string `NA` will
appear in the paths.

```{r}
# Create subdirectory matrix as data frame
(subdir_data <- kwb.file::to_subdir_matrix(paths, result_type = "data.frame"))

# Reconstruct the paths from this data frame
paths_reconstructed <- kwb.pathdict:::data_frame_to_paths(subdir_data)

# Check that the paths have not changed
identical(paths_reconstructed, paths)
```

### to_placeholder(), is_placeholder(), to_dictionary_key()

The function `to_placeholder()` converts a string to a placeholder \<string\>.

```{r}
kwb.pathdict:::to_placeholder("abc")

(placeholders <- kwb.pathdict:::to_placeholder(c("abc", 123)))
```

The function `is_placeholder()` checks whether a string looks like a
\<placeholder\>:

```{r}
kwb.pathdict:::is_placeholder(placeholders)
kwb.pathdict:::is_placeholder("abc")
kwb.pathdict:::is_placeholder("[abc]")
kwb.pathdict:::is_placeholder(c("<abc>", "<def>", "ghi", "<jkl>"))
kwb.pathdict:::is_placeholder("<abc>def")
kwb.pathdict:::is_placeholder("<abc><def>")
kwb.pathdict:::is_placeholder("<>")
```

The function `to_dictionary_key` creates the name for a key in a dictionary,
consisting of a prefix ("p" by default) and a number, given in the hexadecimal 
system:

```{r}
kwb.pathdict:::to_dictionary_key(1:17)
kwb.pathdict:::to_dictionary_key(1:17, leading_zeros = TRUE)
```

### to_dictionary(), use_dictionary()

The function `to_dictionary()` creates a "dictionary" for a given vector of 
input strings. Each unique input string is given a short name by which ich can
be looked up in the returned dictionary. The dictionary is a list with the
unique input strings as values and the assigned short names as keys. The entries
in the dictionary are sorted decreasingly by the "importance" of the 
corresponding input string, i.e. the product of frequency and string length. 

```{r}
dict <- kwb.pathdict:::to_dictionary(dirname(paths))

# Show the dictionary in the form of a data frame
kwb.utils::namedVectorToDataFrame(unlist(dict))
```

Once you have defined a dictionary you can replace the original strings with
placeholders that correspond to the short names in the dictionary:

```{r}
(short_paths <- kwb.pathdict:::use_dictionary(dirname(paths), dict))
```

The short paths can be resolved back to the original paths using the `resolve()`
function from the kwb.utils package:

```{r}
# Resolve the placeholders in the short paths back to the original paths
long_paths <- kwb.utils::resolve(short_paths, dict)

# Check that the original paths were reproduced
identical(long_paths, dirname(paths))
```

### get_dictionary_one_by_one()

```{r}
result <- kwb.pathdict:::get_dictionary_one_by_one(paths, n = 5)
result
```

### Subfolder Frequency Functions

At the start of `get_dictionary_one_by_one()` the function
`get_subdir_frequencies()` is called. This function can be given a vector of
path strings as input. For each possible depth in the path tree all different
sub-paths are determined and ordered by their "importance", i.e. by the product
of frequency and path length in number of characters. By default, only the most
important path per depth is returned:

```{r}
frequencies <- kwb.pathdict:::get_subdir_frequencies(paths = paths, dbg = FALSE)

# Define helper function for output
print_frequencies <- function(f) {
  for (i in seq_along(f)) {
    message("=== depth: ", i)
    print(kwb.utils::namedVectorToDataFrame(f[[i]]))
  }
}

print_frequencies(frequencies)
```

To return all different paths, set `first.only = FALSE`:

```{r}
all_frequencies <- kwb.pathdict:::get_subdir_frequencies(
  paths = paths, first.only = FALSE, dbg = FALSE
)

print_frequencies(all_frequencies)
```

The function `to_frequency_data()` is called next within 
`get_dictionary_one_by_one()`. It converts the list returned by 
`get_subdir_frequencies()` to a data frame.

```{r}
# Convert frequency list to data frame
(frequency_data <- kwb.pathdict:::to_frequency_data(frequencies))
(all_frequency_data <- kwb.pathdict:::to_frequency_data(all_frequencies))
```

The function `rescore_and_reorder_frequency_data()` takes a data frame with
columns `length` and `count` as input. It calculates `score2 = (length -
placeholder_size) * count` and orders the data frame decreasingly by this score.

```{r}
kwb.pathdict:::rescore_and_reorder_frequency_data(
  frequency_data, 
  # Length of placeholder string to be used instead of path
  placeholder_size = nchar("<p1>")
)
```

### get_next_level()

The function `get_next_level()` is a helper function that is called by the
function `compress_one_by_one()`.

```{r}
kwb.pathdict:::get_next_level
```

### get_subdirs_by_frequency()

The function `get_subdirs_by_frequency()` is used by a FAKIN script that is not
yet part of a package. It gets different inputs:

1. `subdirs`: subdirectory matrix
2. `cumid`: cumulative identifier, created with
  `kwb.pathdict:::to_cumulative_id()`
3. `freqinfo`: one-row data frame with columns `depth`, `n.x`, `n.Freq`

```{r}
subdirs <- kwb.file::to_subdir_matrix(paths)
cumid <- kwb.pathdict:::to_cumulative_id(subdirs)
freqinfo <- data.frame(depth = 2, n.x = 10, n.Freq = 10)

kwb.pathdict:::get_subdirs_by_frequency(subdirs, cumid, freqinfo)
```

### Functions that still need to be described

```{r}
kwb.pathdict:::log_result_if
kwb.pathdict:::lookup_in_dictionary
kwb.pathdict:::print_path_frequencies
kwb.pathdict:::replace_subdirs
kwb.pathdict:::starts_with_parts
kwb.pathdict:::update_frequency_data_length
```
