---
title: "Working with File Paths in R Scripts"
author: "Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In many cases R scripts read input files and create output files. The files are
specified by their paths on the file system. Each file path that is hard-coded
in an R script represents a dependency of the script. The script will fail if
files that it accessess are removed or moved on the file system. 

One solution to this problem is not to change paths on the file server. This is
not a good solution. Very often, existing directory structures are not good from
the beginning on. It is good practice to revise existing directory structures
regularly and to optimise them so that they comply with current requirements.

So the aim should be to write the R script in a way that changes in file paths
can be easily adapted in the script. We propose the following to achieve this:

1. Define the paths to all required directories in one place, preferably on top
of the script. 
2. Hold all path definitions in a list of named elements. 
3. Avoid duplication by replacing common parts of paths with placeholders.

### Define example paths

Let's start with an example. Assume that you have revised a script and that you
have found the following hard-coded file paths:

```{r}
paths <- c(
  "//very/long/path/to/the/projects/project-1/wp-1/input/file 1.csv",
  "//very/long/path/to/the/projects/project-1/wp-1/input/file-2.csv",
  "//very/long/path/to/the/projects/project-1/wp-1/analysis/summary.pdf",
  "//very/long/path/to/the/projects/project-1/wp 2/input/köpenick_dirty.csv",
  "//very/long/path/to/the/projects/project-1/wp 2/output/koepenick_clean.csv",
  "//very/long/path/to/the/projects/project-2/Daten/file-1.csv",
  "//very/long/path/to/the/projects/project-2/Grafiken/file 1.png",
  "//very/long/path/to/the/projects/project-2/Berichte/bericht-1.doc",
  "//very/long/path/to/the/projects/project-2/Berichte/bericht-2.doc"
)
``` 

We propose to give a symbolic name to each of the different directories that
are involved. These are:

```{r}
unique(dirname(paths))
```

### Define directory paths in one place

Define these directory paths in a list (`dirs_1` in the following) and give each 
directory a short but meaningful name:

```{r}
dirs_1 <- list(
  p1.1_input = "//very/long/path/to/the/projects/project-1/wp-1/input"   ,
  p1.1_analysis = "//very/long/path/to/the/projects/project-1/wp-1/analysis",
  p1.2_input = "//very/long/path/to/the/projects/project-1/wp 2/input",
  p1.2_output = "//very/long/path/to/the/projects/project-1/wp 2/output",
  p2_data = "//very/long/path/to/the/projects/project-2/Daten",
  p2_images = "//very/long/path/to/the/projects/project-2/Grafiken",
  p2_reports = "//very/long/path/to/the/projects/project-2/Berichte"
)
```

Whenever you use a file in the script, create the full file path by calling 
`file.path()`, as in the following example:

```{r}
# Instead of:
# "//very/long/path/to/the/projects/project-1/wp 2/output/koepenick_clean.csv"

# Compose the full path from the directory path and the file name:
file.path(dirs_1$p1.2_output, "koepenick_clean.csv")
```

We prefer to do it this way instead of defining the full file path in a list,
e.g. `files`, and then accessing it with e.g. 
`files$p1.2_output_koepenick_clean`, for two reasons:

1. We would need to find a meaningful name for the list element and we would
most probably end up in a name that looks very similar to the file name itself.
2. If only the file name changes, we need to change it only in one place (where 
we call `file.path`). Otherwise we needed to change the list `files` as well.

### Replace common parts of paths with placeholders

The list of directory paths contains a lot of duplication. This not only makes 
it difficult to read but also hinders the adaptation of paths in case that the 
path to a common base directory changes. If, for example the path

```
//very/long/path/to/the/projects
```

moved to somewhere else, we would need to change it seven times in the list.
This could be tedious and could result in errors by missing out a replacement.
Therefore, we suggest to define common base paths on their own in the list and
to use placeholders of the form `<name>` in the paths where they occur. This
could look as follows:

```{r}
dirs_2 <- list(
  projects = "//very/long/path/to/the/projects",
  p1.1_input = "<projects>/project-1/wp-1/input",
  p1.1_analysis = "<projects>/project-1/wp-1/analysis",
  p1.2_input = "<projects>/project-1/wp 2/input",
  p1.2_output = "<projects>/project-1/wp 2/output",
  p2_data = "<projects>/project-2/Daten",
  p2_images = "<projects>/project-2/Grafiken",
  p2_reports = "<projects>/project-2/Berichte"
)
```

We could go one step futher and define list entries for each project (`p1` and
`p2` in the following) as well:

```{r}
dirs_3 <- list(
  projects = "//very/long/path/to/the/projects",
  p1 = "<projects>/project-1",
  p2 = "<projects>/project-2",
  p1.1_input = "<p1>/wp-1/input",
  p1.1_analysis = "<p1>/wp-1/analysis",
  p1.2_input = "<p1>/wp 2/input",
  p1.2_output = "<p1>/wp 2/output",
  p2_data = "<p2>/Daten",
  p2_images = "<p2>/Grafiken",
  p2_reports = "<p2>/Berichte"
)
```

Finally, we could even define entries for each directory representing a work 
package (wp) of project 1 (`p1.1` and `p1.2` in the following):

```{r}
dirs_4 <- list(
  projects = "//very/long/path/to/the/projects",
  p1 = "<projects>/project-1",
  p2 = "<projects>/project-2",
  p1.1 = "<p1>/wp-1",
  p1.2 = "<p1>/wp-2",
  p1.1_input = "<p1.1>/input",
  p1.1_analysis = "<p1.1>/analysis",
  p1.2_input = "<p1.2>/input",
  p1.2_output = "<p1.2>/output",
  p2_data = "<p2>/Daten",
  p2_images = "<p2>/Grafiken",
  p2_reports = "<p2>/Berichte"
)
```

The above list `dirs_4` may look a bit over complicated but it represents a path
definition that is free of redundancies. This has the advantage that, if any
directory path changes, only one entry of the list needs to be changed. Also, 
this condensed form of the path definition better reveals the directory
structure than did the very first version `dirs_1` that we defined above.

Of course, we cannot use this version containing `<placeholders>` directly in
calls to `file.path` as we did with `dirs_1`. We need a means to create the full 
paths as defined in `dirs_1` back from the condensed path definitions in 
`dirs_4`. 

This is what the function `resolve()` from our package kwb.utils does. You may
use this function directly around the list assignment as in the following:

```{r}
dirs_5 <- kwb.utils::resolve(list(
  projects = "//very/long/path/to/the/projects",
  p1 = "<projects>/project-1",
  p2 = "<projects>/project-2",
  p1.1 = "<p1>/wp-1",
  p1.2 = "<p1>/wp 2",
  p1.1_input = "<p1.1>/input",
  p1.1_analysis = "<p1.1>/analysis",
  p1.2_input = "<p1.2>/input",
  p1.2_output = "<p1.2>/output",
  p2_data = "<p2>/Daten",
  p2_images = "<p2>/Grafiken",
  p2_reports = "<p2>/Berichte"
))
```

The `resolve()` function replaces (recursively) all placeholders by their 
corresponding values. As a result, the list now contains the full paths, just
as they were defined in the first version `dirs_1` of the list. We prove this by
the following comparison:

```{r}
identical(dirs_5[names(dirs_1)], dirs_1)
```

